---
title: "The Autocorrelation Function"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Creating a purely random time series with a normal distribution 
- with the use of the rnorm() routine

```{r}
purely_random_process=ts(rnorm(100)) # 100 data points from std normal distribution that is a ts objecct
(acf(purely_random_process, type='covariance'))
```

- We assume weak stattionarity
  - No systematic change in mean or variation
  - No periodic fluctuations
  
The acf() routine plots autocorrelation cofficients at differnt lags: correlogram
since it starts at 1 r0 = c0/c0 = 1

The resulting plot is the autocorrelation cofficients at different lags. 

```{r}
( acf(purely_random_process, main='Correlogram of a Purely Random Porcess') )
```

R0 always starts at 1 and then we see that, then we see there isnt much correlation between the other lags, and based on our selection of a normal distribution and therefore we dont expect any correlation between the lags. 

The dashed lines show the significance leves and none of the lag spikes crossing the dashed lines indiciate that there isnt any correlation between our lags. 







