---
title: "Visualizing and Describing a Time Series, time plots, autocovariance, and auto correction"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Johnson Johnson Quarterly Earnings Analysis of (jj) data set

- US company Johnson and Johnson
- Data set contains 84 quarterly earnings
- Q1 1960 - Q4 1980

```{r}
require(astsa)
plot(jj, type='o', main='Johnson & Johnson quarterly earnings per share', ylab='Earnings', xlab= 'quarters')
```

- There is also a general upwards trend as the quarters increase so does EPS
- At the beginning there isn't much variation but as the quarters increase there is alot more fluctuation 
- This violates the Stationary principle
- There is also apparent seasonal changes on the trend


## Pneumonia and Influenza Deaths in the US Analysis of (flu) data set
- Monthly pneumonia and influenza deaths per 10 000 people 
- 11 years
- from 1968 to 1978

```{r}
plot(flu, main="Monthly Pneumonia and Influenza Deaths in the US", ylab= 'Number of Deaths per 10 000 people', xlab='Months')
```

- Very apparant seasonality, peak after every year in the dataset maybe in the summer
- not a stationary time series
- Overall trend is decreasing over the yeras

## Land-Ocean temperature deviations; Analysis of the (globtemp) dataset

- Global mean land-ocean temp deviations
- deviations from the base period 1951 - 1980
- Measured in degrees celcius
- For the years 1880 - 2015 
- From NASA

```{r}
plot(globtemp, main='Global mean land-ocean deviations from the average temp of 1951-1980', ylab='Temperature deviations', type='o', xlab='Years')
```


- From this timeplot there is is an apparnt increase in deviations
- There is also seasonality in the trend --> maybe due to seaons?

## Variable Star; analysis of the (star) dataset
- The magnitude of a star taken at midnight
- For 600 consecutive days
- The data are from the "The Calculus of Observations..."

```{r}
plot(star, main="The magniture of a star taken at mightnight for 600 consective days", ylab = 'Magniture', xlab='Days')
```

- There is apparant seasonality and perodicity in this data set.